Macroscopic Salt Rejection through Electrostatically Gated Nanoporous Graphene

### 2019-08-20 paper/paper.tex
Change after final decision from NL

\documentclass[journal=nalefd,email=true, hyperref=true, keywords=false]{achemso}
\usepackage[utf8]{inputenc}
\usepackage{graphicx}
\usepackage{tabularx}
\usepackage{amsmath}
\usepackage{times}
\usepackage{todonotes}
\DeclareCaptionLabelSeparator{bar}{  \textbar\quad}
\usepackage[labelsep=bar, labelfont=bf]{caption}
\newcommand{\Fig}{Fig.}
% \usepackage{lineno}
\usepackage{units}
\usepackage{txfonts}
\usepackage{xr-hyper}
\externaldocument[SI-]{SI}

@@

\keywords{graphene, nanopore, patterning, salt rejection, gating, finite element modeling, quantum capacitance}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Main text
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\clearpage{}
% \linenumbers
\section{Introduction}
\label{sec:intro}
A central objective in the emerging field of nanofluidics is to
control transport phenomena at nanometer scale to explore new
technological opportunities. The reduction of characteristic
dimensions in nanofluidics gives rise to substantially different
transport properties in comparison to their bulk counterparts
\cite{Schoch_2008}. In nanopores, for example, the effect of
surface-mediated transport becomes dominant when the pore size is
comparable to the characteristic length of diffusion, enabling new
applications such as ionic diodes
\cite{Karnik_2007,siwy2002fabrication,vlassiouk2007nanofluidic},
field-effect transistors \cite{Nam_2009}, desalination
\cite{Heiranian_2015}, and nanopore-based DNA sequencing
\cite{Heerema_2016,Garaj_2013}. Porous, atomically thin
two-dimensional (2D) materials, such as graphene, are emerging as
ideal membrane materials due to the ultimate permeation
\cite{Suk_2010,Jiang_2009,Celebi_2014,Koenig_2012,Drahushuk_2012}. Early
research has suggested that using a sheet of nanoporous graphene can
act as a desalination membrane by having the pore sizes smaller than
the hydrated radii ($\sim$1 nm), thereby hindering ion passage without
losing water permeability
\cite{Cohen_Tanugi_2012,Suk_2014,Cohen_Tanugi_2014,Cohen_Tanugi_2015,O_Hern_2014,O_Hern_2015,Surwade_2015,Walker_2017,Ghosh_2018}.
Nevertheless, in practice, the fabrication of subnanometer porous 2D
membranes over a large area with atomic precision remains technically
challenging, { limiting the effective separation or demanding
  leak-sealing mechanism such as interfacial polymerization}
\cite{Suk_2014,Rollings_2016,O_Hern_2012,Wang_2017}. Moreover, a small
portion of excessively large pores or tears can significantly diminish
salt rejection. Regarding these issues, we have previously developed a
highly-scalable technique to fabricate uniformly-distributed nanopores
(sub-20 to 50 nm) over wafer-scale graphene samples
\cite{Choi_2018}. A major step towards its real-world desalination and
separation applications, would be to develop methodology that allow
appreciable selectivity of nanoporous 2D membrane with pore size even
larger than the hydration radii.
% Therefore, it is critical to develop methodology that
% allow appreciable selectivity of nanoporous 2D membrane even with
% large pore size.
A number of recent reports have explored the
effect of surface charges on the selectivity of ion transport through
graphene nanopores, as a consequence of the electrostatic interactions
between solvated ions and the charged functional groups attached on
graphene, indicating that even large pores ($>$ 30nm) can be ion
selective \cite{Rollings_2016,Surwade_2014}. However, it remains
controversial that if the surface charges can result in a degree of
salt rejection, as the surface charge and potential cannot be
precisely quantified and are sensitive to the surroundings as well as
the history of chemical treatment \cite{Li_2008}. There remains a lack
of fundamental understanding of the interactions between the charged
2D nanopores and the electrolyte solution, { especially in regard of controlled ion transport modulation}.

In this report, we
investigate electrolyte permeation through graphene nanopores that are
significantly larger than the hydration radii. We model the surface
potential by taking into account the elementary electronic
properties of graphene and demonstrate that a high degree of salt
rejection could be achieved by both pore sizing and electrostatic gating.

@@

The above mechanistic findings, nevertheless, do not fully clarify the
experimentally observed reduction of diffusive flux upon
gating. Indeed, the same physical mechanism also governs 
the 
{ diode-like ionic
transport through an IFET  \cite{Nam_2009,Lee_2015,Feng_2016}
or a nanoporous metal membrane \cite{mccurry2017electrolyte}},
in which a nonzero electric potential at
the nanopore center usually results in an increase of ionic
conductivity. To this end, we
further increased the graphene surface potential $\psi_{\mathrm{G}}$
considering the same system (details see Supplementary Section
\ref{SI-sec:conc}). Intriguingly, by increasing
$\psi_{\mathrm{G}}$ larger than 400 mV, the ionic flux starts to
increase, reversing the trend observed at the low $\psi_{\mathrm{G}}$
regime, due to (i) the drift flux become dominant over diffusion flux
and (ii) the concentration near the pore edge (preferred path) greatly
increases (details see Supplementary Section \ref{SI-sec:conc},
Supplementary Figs. \ref{SI-fig:reverse} and
\ref{SI-fig:large-V}). This level of $\psi_{\mathrm{G}}$ cannot be
reached experimentally in graphene membrane, as this requires a
$V_{\mathrm{G}}$ larger than 2.0 V, which triggers the electrochemical
reactions \cite{Toh_2011}. On the other hand, in an IFET, the electric
potential on the pore wall is considerably higher, equivalent to
$V_{\mathrm{G}}$ due to the metallic nature of gate electrode (having
an infinitely-large quantum capacitance). We therefore conclude that
the quantum capacitance-induced nonlinear damping in the surface
potential results in the observed salt rejection.  Following the above
discussions, we further investigate the physical limits for the biased
graphene nanopores. Specifically, the salt rejection characteristics
are controlled by: (i) the overlap of EDL inside the nanopore,
essentially controlled by two length scales of $\lambda_{\mathrm{D}}$
and $r_{\mathrm{G}}$, and (ii) the graphene surface potential
$\psi_{\mathrm{G}}$ controlled by $V_{\mathrm{G}}$ following
Eq. \eqref{eq:Vg}. To this end, we calculate $\psi_{\mathrm{G}}$ as a
function of $\lambda_{\mathrm{D}} / r_{\mathrm{G}}$ and
$V_{\mathrm{G}}$ (\Fig{} \ref{fig:5}a). Clearly, when the bulk
concentration $c_{0}$ increases, a higher $V_{\mathrm{G}}$ is required
to reach the same level of $\psi_{\mathrm{G}}$. Consequently, as
revealed in \Fig{} \ref{fig:5}b, the calculated salt rejection factor
$\xi$ increases with both $\lambda_{\mathrm{D}} / r_{\mathrm{G}}$ and
$V_{\mathrm{G}}$, in line with the experimentally observed
$\xi - c_{0}$ dependence. We also find the overall trends for
$\psi_{\mathrm{G}}$ and $\xi$ with respect to
$\lambda_{\mathrm{D}}/r_{\mathrm{G}}$ and $V_{\mathrm{G}}$ are very
similar. More importantly, based on our theoretical prediction, with
$V_{\mathrm{G}}$ = 1.25 V and $\lambda_{\mathrm{D}} / r_{\mathrm{G}}$
= { 2.0},
a high value of $\xi$ up to $\sim$0.95 can be achieved.  We
notice that although the above analysis suggests a salt rejection up
to 1 may be achived by further reducing the pore size (i.e. increasing
the $\lambda_{\mathrm{D}}/r_{\mathrm{G}}$ ratio), our theory, derived
from on the macroscopic transport equations by treating ions as
zero-volume charges, may not be valid when $r_{\mathrm{G}}<$ 5
nm\cite{Jain_2015}. In addition, the specific ion-graphene
interactions \cite{Rollings_2016} are not taken into account. Advanced
theoretical frameworks that bridge continuum and molecular models are
required to elucidate electrolyte transport approaching the small pore
limit. 
{
We notice that the calculations in Figs. \ref{fig:4} and \ref{fig:5} 
are based on the setup  of ionic transport through single nanopore.
In order to mimic the experimental conditions, 
calculations were carried out by considering the pore size distribution in Fig. \ref{fig:1}c inset. 
Under the assumption that the interpore distance is significantly larger than the Debye length,
the overall salt rejection $\xi$ is
estimated by the linear combination of the salt rejection 
$\xi_{r_i}$ 
from individual pores with radius $r_i$ given by:
\label{eq:rejection-comb}
\xi = \sum_{r_{i}} \xi_{r_{i}} w_{r_{i}}
$w_{r_{i}} = x_{r_{i}} r_{i}^{2} / \sum_{r_{i}} x_{r_{i}} r_{i}^{2} $
is the contribution factor from pores with radius $r_{i}$, and
$x_{r_{i}}$ is the fraction of pores with radius $r_{i}$ (details see
Supplementary Section \ref{SI-sec:pore-dist}). Combining
Eq. \ref{eq:rejection-comb} and \Fig{} \ref{fig:5}b, the calculated
salt rejection is reduced to $\sim{}$ 80\% of that from the
single-pore calculation of a 20 nm nanopore when $c_{0}$ = 0.1 mM
(Supplementary \Fig{} \ref{SI-fig:simple-rect-pore}). Consider the
existence of large pores with diameter up to 60 nm, such degree of
salt rejection is still appreciable, which is ascribed to the
nonlinear dependency of $\xi$ on
$\lambda_{\mathrm{D}} / r_{\mathrm{G}}$ and $V_{\mathrm{G}}$ (details
see Supplementary Section \ref{SI-sec:pore-dist}).}  From our
analysis, it is also clear reducing the average pore size is critical
to overcome the low salt rejection at higher concentrations, which may
hopefully be achieved with more advanced fabrication techniques.

@@

Finally, we examine other salt species and compare the degree of salt
rejection between experiments and simulations. \Fig{} \ref{fig:6}a
presents the measured salt rejection $\xi$ as a function of
$V_{\mathrm{G}}$ at $c_{0}$ = 0.1 mM, for the seven salt species
considered here, including NaCl, LiCl, KCl, MgSO$_{4}$, CaCl$_{2}$,
K$_{2}$SO$_{4}$, and K$_{3}$[Fe(CN)$_{6}$] (details see Supplementary
Section \ref{SI-sec:salts}). Similar to the KCl system discussed
earlier, the salt rejection is more pronounced in the positive
$V_{\mathrm{G}}$ regime. 
% { The negative $\eta$ values, observed for certain salts or applied $V_{\mathrm{G}}$, respectively, would actually
% indicate that the flow in enhanced upon biasing the membrane. While such behavior is not uncommon in gating systems, we cannot conclude unambiguously if the effect is significant or rather within the measurement uncertainty of the system. Furthermore, as the definition of $\xi$ based on $\delta$ is only valid for positive $\eta$, the discussion is hence limited to $V_{\mathrm{G}}$ $>$ 0.}
The degree of salt rejection deceases with
the salt valance; for example, the measured $\xi$ values at
$V_{\mathrm{G}}$ = 1.25 V decreases from $>$ 0.5 for monovalent NaCl
to -0.04 for multivalent K$_{3}$[Fe(CN)$_{6}$]. This observation is
endorsed by our simulations, in which we calculate $\xi$ as a function
of $V_{\mathrm{G}}$ using the same simulation setups (\Fig{}
\ref{fig:6}b). 
{
The effect of pore size distribution is explicitly
considered using the FEM simulation results with different pore sizes
(Supplementary \Fig{} \ref{SI-fig:xi-salts-pore})} 
The calculated $\xi_{\mathrm{max}}$ values are
quantitatively compared with experiments (\Fig{} \ref{fig:6}c), and
the same trend is observed. Indeed, because the Debye length
$\lambda_{\mathrm{D}}$ decreases with the salt valence, the graphene
surface potential appears to decrease in the multivalent salt systems,
thereby reducing $\xi$ (see \Fig{} \ref{fig:5}b).
{
The slightly higher $\xi$ values of LiCl and NaCl compared with KCl observed in the experiments, 
are also supported by the FEM simulation results in \Fig{} \ref{fig:6}b, 
which we attribute to the lower diffusivities of Li$^{+}$ and Na$^{+}$ ions (see Supplementary Table \ref{SI-tab:diff})
}.
The calculated and
experimental $\xi_{\mathrm{max}}$ as a function of
$\lambda_{\mathrm{D}}$ are consequently plotted, for the seven salt
species considered here (\Fig{} \ref{fig:6}d). 
% Similar to the
% experimental observations, slightly higher $\xi$ of salts
% with same valence (e.g. LiCl, NaCl and KCl) is also predicted from the
% simulations due to the variation in ionic mobilities. 
We notice that
without any fitting parameters, our theory can predict the
experimental values reasonably well.  In addition, the
$\xi_{\mathrm{max}}-\lambda_{\mathrm{D}}$ dependence is approximately
linear, suggesting that the effect of Debye length is dominant over
the salt diffusivity. This finding also explains the concentration
dependence of $\xi_{\mathrm{max}}$ observed in \Fig{} \ref{fig:2}d, as
$\lambda_{\mathrm{D}} \propto c_{0}^{-0.5}$. Further including the
specific interaction of ionic species with the graphene membrane
{- such as cation-$\pi$ interacation \cite{shi2013ion}} - may give even better agreement with experimental
results \cite{Ghosh_2018}. More interestingly, based on our theoretical analysis, we
predict that a very high degree of salt rejection ($\xi>$ 0.99) can be
already achieved when $\lambda / r_{\mathrm{G}}>$3. In other words,
for example, when the Debye length is 20 nm, one can suppress 99 \% of
the diffusive flux by further reducing the pore radius down to
$\sim$5-6 nm, which still remains larger than the hydrated radii. In
combination with the development of nanoscale etching techniques, we
believe that the physical picture presented here makes one step closer
to large-scale production of graphene membranes.

@@

\begin{acknowledgement}
  R.M.W. acknowledges Dr. Kyoungjun Choi from ETH Zurich for the measurement of the electron mobility 
  of patterned graphene and is grateful for the support from the Binnig and Rohrer
  Nanotechnology Center of ETH Zurich and IBM Zurich. T.T. and
  C.J.S. are grateful for the financial support from the ETH startup
  funding. 
\end{acknowledgement}

@@

@@

\begin{figure}[H]
  \centering
  \includegraphics[width=0.95\linewidth]{img/fig2.pdf}
  \caption{\textbf{The experimental evidence of salt rejection.}
    \textbf{a}.An example of measured conductivity in LCR as a
    function of time. The slope of conductivity-time curve when
    $V_{\mathrm{G}}>0$ is interpolated, is smaller than that without
    gating. \textbf{b}. Molar diffusive flux through bare PCTE (blue
    bars: experimental; red bars: model values) and PG+PCTE (green
    bars) membranes for different salts. \textbf{c}. Experimentally
    measured salt rejection factor $\xi$ as a function of
    $V_{\mathrm{G}}$ for 0.1 mM KCl solution in HCR, showing an
    asymmetric response, with a higher degree of salt rejection for
    $V_{\mathrm{G}}>0$. \textbf{d}. $\xi_{\mathrm{max}}$ as a function
    of KCl concentration, fitted by a power law relation, suggesting
    the effect of Debye length on the observed salt rejection, since
    $\lambda_{\mathrm{D}} \propto c_{0}^{-\frac{1}{2}}$.}
  \label{fig:2} 
\end{figure}

@@

>>>>

\documentclass[journal=nalefd,email=true, hyperref=true, keywords=false]{achemso}
\usepackage[utf8]{inputenc}
\usepackage{graphicx}
\usepackage{tabularx}
\usepackage{amsmath}
\usepackage{times}
\usepackage{todonotes}
\DeclareCaptionLabelSeparator{bar}{  \textbar\quad}
\usepackage[labelsep=bar, labelfont=bf]{caption}
\newcommand{\Fig}{Fig.}
%%Uncomment to disable color changes
\renewcommand*{\textcolor}[2]{#2}
% \usepackage{lineno}
\usepackage{units}
\usepackage{txfonts}
\usepackage{xr-hyper}
\externaldocument[SI-]{SI}

@@

% \keywords{graphene, nanopore, patterning, salt rejection, gating, finite element modeling, quantum capacitance}

\keywords{\textcolor{blue}{nanoporous graphene}, salt rejection, gating, finite element modeling, quantum capacitance}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Main text
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\clearpage{}
% \linenumbers
\section{Introduction}
\label{sec:intro}
A central objective in the emerging field of nanofluidics is to
control transport phenomena at nanometer scale to explore new
technological opportunities. The reduction of characteristic
dimensions in nanofluidics gives rise to substantially different
transport properties in comparison to their bulk counterparts
\cite{Schoch_2008}. In nanopores, for example, the effect of
surface-mediated transport becomes dominant when the pore size is
comparable to the characteristic length of diffusion, enabling new
applications such as ionic diodes
\cite{Karnik_2007,siwy2002fabrication,vlassiouk2007nanofluidic},
field-effect transistors \cite{Nam_2009}, desalination
\cite{Heiranian_2015}, and nanopore-based DNA sequencing
\cite{Heerema_2016,Garaj_2013}. Porous, atomically thin
two-dimensional (2D) materials, such as graphene, are emerging as
ideal membrane materials due to the ultimate permeation
\cite{Suk_2010,Jiang_2009,Celebi_2014,Koenig_2012,Drahushuk_2012}. Early
research has suggested that using a sheet of nanoporous graphene can
act as a desalination membrane by having the pore sizes smaller than
the hydrated radii ($\sim$1 nm), thereby hindering ion passage without
losing water permeability
\cite{Cohen_Tanugi_2012,Suk_2014,Cohen_Tanugi_2014,Cohen_Tanugi_2015,O_Hern_2014,O_Hern_2015,Surwade_2015,Walker_2017,Ghosh_2018}.
Nevertheless, in practice, the fabrication of subnanometer porous 2D
membranes over a large area with atomic precision remains technically
challenging, limiting the effective separation or demanding
  leak-sealing mechanism such as interfacial polymerization
\cite{Suk_2014,Rollings_2016,O_Hern_2012,Wang_2017}. Moreover, a small
portion of excessively large pores or tears can significantly diminish
salt rejection. Regarding these issues, we have previously developed a
highly-scalable technique to fabricate uniformly-distributed nanopores
(sub-20 to 50 nm) over wafer-scale graphene samples
\cite{Choi_2018}. A major step towards its real-world desalination and
separation applications, would be to develop methodology that allow
appreciable selectivity of nanoporous 2D membrane with pore size even
larger than the hydration radii.
% Therefore, it is critical to develop methodology that
% allow appreciable selectivity of nanoporous 2D membrane even with
% large pore size.
A number of recent reports have explored the effect of surface charges
on the selectivity of ion transport through graphene nanopores, as a
consequence of the electrostatic interactions between solvated ions
and the charged functional groups attached on graphene, indicating
that even large pores ($>$ 30nm) can be ion selective
\cite{Rollings_2016,Surwade_2014}. However, it remains controversial
that if the surface charges can result in a degree of salt rejection,
as the surface charge and potential cannot be precisely quantified and
are sensitive to the surroundings as well as the history of chemical
treatment \cite{Li_2008}. There remains a lack of fundamental
understanding of the interactions between the charged 2D nanopores and
the electrolyte solution, especially in regard of controlled ion
transport modulation.

% \textcolor{blue}{In this report, we investigate electrolyte permeation
  % through cm$^2$ scale graphene membranes having pores that are
  % $<$\unit[50]{nm}. We find, that an applied voltage bias can
  % effectively modulate the ion transport through these membranes,
  % despite having pores significantly larger than the hydration radii
  % of the ions.  We simulate the surface potential, taking into account
  % the elementary electronic properties of graphene, allowing to model
  % and confirm that a high degree of salt rejection could be achieved
  % by both pore sizing and electrostatic gating. Our analysis reveals a
  % large impact of atomistic effects, such as quantum capacitance, on
  % the ionic transport through gated porous 2D materials, providing new
  % insights into such phenomena.}

\textcolor{blue}{ In this report, we investigate electrolyte
  permeation through centimeter scale graphene membranes having pores
  that are in the order of $\sim{}$\unit[20]{nm}. We find that
  electrostatic gating can effectively reduce ion transport through
  these membranes, despite the pore size is significantly larger than
  the hydration radii of the ions. In the literature, it was only
  observed in the membranes made by mesoporous carbon
  \cite{Surwade_2014}. For the first time, we model the surface
  potential by taking into account the elementary electronic
  properties of graphene and confirm that a high degree of salt
  rejection could be achieved by both pore sizing and electrostatic
  gating. }

@@

The above mechanistic findings, nevertheless, do not fully clarify the
experimentally observed reduction of diffusive flux upon
gating. Indeed, the same physical mechanism also governs the {
  diode-like ionic transport through an IFET
  \cite{Nam_2009,Lee_2015,Feng_2016} or a nanoporous metal membrane
  \cite{mccurry2017electrolyte}}, in which a nonzero electric
potential at the nanopore center usually results in an increase of
ionic conductivity. To this end, we further increased the graphene
surface potential $\psi_{\mathrm{G}}$ considering the same system
(details see Supplementary Section \ref{SI-sec:conc}). Intriguingly,
by increasing $\psi_{\mathrm{G}}$ larger than 400 mV, the ionic flux
starts to increase, reversing the trend observed at the low
$\psi_{\mathrm{G}}$ regime, due to (i) the drift flux become dominant
over diffusion flux and (ii) the concentration near the pore edge
(preferred path) greatly increases (details see Supplementary Section
\ref{SI-sec:conc}, Supplementary Figs. \ref{SI-fig:reverse} and
\ref{SI-fig:large-V}). This level of $\psi_{\mathrm{G}}$ cannot be
reached experimentally in graphene membrane, as this requires a
$V_{\mathrm{G}}$ larger than 2.0 V, which triggers the electrochemical
reactions \cite{Toh_2011}. On the other hand, in an IFET, the electric
potential on the pore wall is considerably higher, equivalent to
$V_{\mathrm{G}}$ due to the metallic nature of gate electrode (having
an infinitely-large quantum capacitance). We therefore conclude that
the quantum capacitance-induced nonlinear damping in the surface
potential results in the observed salt rejection.  Following the above
discussions, we further investigate the physical limits for the biased
graphene nanopores. Specifically, the salt rejection characteristics
are controlled by: (i) the overlap of EDL inside the nanopore,
essentially controlled by two length scales of $\lambda_{\mathrm{D}}$
and $r_{\mathrm{G}}$, and (ii) the graphene surface potential
$\psi_{\mathrm{G}}$ controlled by $V_{\mathrm{G}}$ following
Eq. \eqref{eq:Vg}. To this end, we calculate $\psi_{\mathrm{G}}$ as a
function of $\lambda_{\mathrm{D}} / r_{\mathrm{G}}$ and
$V_{\mathrm{G}}$ (\Fig{} \ref{fig:5}a). Clearly, when the bulk
concentration $c_{0}$ increases, a higher $V_{\mathrm{G}}$ is required
to reach the same level of $\psi_{\mathrm{G}}$. Consequently, as
revealed in \Fig{} \ref{fig:5}b, the calculated salt rejection factor
$\xi$ increases with both $\lambda_{\mathrm{D}} / r_{\mathrm{G}}$ and
$V_{\mathrm{G}}$, in line with the experimentally observed
$\xi - c_{0}$ dependence. We also find the overall trends for
$\psi_{\mathrm{G}}$ and $\xi$ with respect to
$\lambda_{\mathrm{D}}/r_{\mathrm{G}}$ and $V_{\mathrm{G}}$ are very
similar. More importantly, based on our theoretical prediction, with
$V_{\mathrm{G}}$ = 1.25 V and $\lambda_{\mathrm{D}} / r_{\mathrm{G}}$
= { 2.0}, a high value of $\xi$ up to $\sim$0.95 can be achieved.  We
notice that although the above analysis suggests a salt rejection up
to 1 may be achived by further reducing the pore size (i.e. increasing
the $\lambda_{\mathrm{D}}/r_{\mathrm{G}}$ ratio), our theory, derived
from on the macroscopic transport equations by treating ions as
zero-volume charges, may not be valid when $r_{\mathrm{G}}<$ 5
nm\cite{Jain_2015}. \textcolor{blue}{ Furthermore, while the proposed
  simplified model can nicely describe the experimental results for
  the relatively large pore size investigated, such that the
  electrochemical potential-driven diffusion pathway is dominant,
  selective adsorption caused by unique chemical nature of pore edges
  may be significant in the small-pore systems. Therefore, advanced
  theoretical frameworks that bridge continuum and molecular models
  are required to elucidate the role of surface adsorption in
  electrolyte transport through approaching the small pore limit. }

% \textcolor{blue}{Furthermore, the specific adsorption of ions
  % \cite{bard_electrochemical_1980} on porous graphene surface with
 % on highly-charged sites \cite{Rollings_2016} is ignored in the current
  % model due to (i) low intrinsic charge density of porous graphene
  % fabricated using the BCP-patterning method and (ii) the surface charge
  % on graphene being effectively dominated by gating (Supplementary
  % Section \ref{SI-sec:charge-dens}). A full kinetic description of
  % such interactions should, however, be included in the model in case 
  % highly-adsorbing species such as fluoride or organic ions are
  % involved \cite{He_2012}.}

{
We notice that the calculations in Figs. \ref{fig:4} and \ref{fig:5} 
are based on the setup  of ionic transport through single nanopore.
In order to mimic the experimental conditions, 
calculations were carried out by considering the pore size distribution in Fig. \ref{fig:1}c inset. 
Under the assumption that the interpore distance is significantly larger than the Debye length,
the overall salt rejection $\xi$ is
estimated by the linear combination of the salt rejection 
$\xi_{r_i}$ 
from individual pores with radius $r_i$ given by:
\begin{equation}
\label{eq:rejection-comb}
\xi = \sum_{r_{i}} \xi_{r_{i}} w_{r_{i}}
\end{equation}
where
$w_{r_{i}} = x_{r_{i}} r_{i}^{2} / \sum_{r_{i}} x_{r_{i}} r_{i}^{2} $
is the contribution factor from pores with radius $r_{i}$, and
$x_{r_{i}}$ is the fraction of pores with radius $r_{i}$ (details see
Supplementary Section \ref{SI-sec:pore-dist}). Combining
Eq. \ref{eq:rejection-comb} and \Fig{} \ref{fig:5}b, the calculated
salt rejection is reduced to $\sim{}$ 80\% of that from the
single-pore calculation of a 20 nm nanopore when $c_{0}$ = 0.1 mM
(Supplementary \Fig{} \ref{SI-fig:simple-rect-pore}). Consider the
existence of large pores with diameter up to 60 nm, such degree of
salt rejection is still appreciable, which is ascribed to the
nonlinear dependency of $\xi$ on
$\lambda_{\mathrm{D}} / r_{\mathrm{G}}$ and $V_{\mathrm{G}}$ (details
see Supplementary Section \ref{SI-sec:pore-dist}).}  From our
analysis, it is also clear reducing the average pore size is critical
to overcome the low salt rejection at higher concentrations, which may
hopefully be achieved with more advanced fabrication techniques.

@@

Finally, we examine other salt species and compare the degree of salt
rejection between experiments and simulations. \Fig{} \ref{fig:6}a
presents the measured salt rejection $\xi$ as a function of
$V_{\mathrm{G}}$ at $c_{0}$ = 0.1 mM, for the seven salt species
considered here, including NaCl, LiCl, KCl, MgSO$_{4}$, CaCl$_{2}$,
K$_{2}$SO$_{4}$, and K$_{3}$[Fe(CN)$_{6}$] (details see Supplementary
Section \ref{SI-sec:salts}). Similar to the KCl system discussed
earlier, the salt rejection is more pronounced in the positive
$V_{\mathrm{G}}$ regime. 
% { The negative $\eta$ values, observed for certain salts or applied $V_{\mathrm{G}}$, respectively, would actually
% indicate that the flow in enhanced upon biasing the membrane. While such behavior is not uncommon in gating systems, we cannot conclude unambiguously if the effect is significant or rather within the measurement uncertainty of the system. Furthermore, as the definition of $\xi$ based on $\delta$ is only valid for positive $\eta$, the discussion is hence limited to $V_{\mathrm{G}}$ $>$ 0.}
The degree of salt rejection deceases with the salt valance; for
example, the measured $\xi$ values at $V_{\mathrm{G}}$ = 1.25 V
decreases from $>$ 0.5 for monovalent NaCl to -0.04 for multivalent
K$_{3}$[Fe(CN)$_{6}$]. This observation is endorsed by our
simulations, in which we calculate $\xi$ as a function of
$V_{\mathrm{G}}$ using the same simulation setups (\Fig{}
\ref{fig:6}b).  { The effect of pore size distribution is explicitly
  considered using the FEM simulation results with different pore
  sizes (Supplementary \Fig{} \ref{SI-fig:xi-salts-pore})} The
calculated $\xi_{\mathrm{max}}$ values are quantitatively compared
with experiments (\Fig{} \ref{fig:6}c), and the same trend is
observed. Indeed, because the Debye length $\lambda_{\mathrm{D}}$
decreases with the salt valence, the graphene surface potential
appears to decrease in the multivalent salt systems, thereby reducing
$\xi$ (see \Fig{} \ref{fig:5}b).  { The slightly higher $\xi$ values
  of LiCl and NaCl compared with KCl observed in the experiments, are
  also supported by the FEM simulation results in \Fig{} \ref{fig:6}b,
  which we attribute to the lower diffusivities of Li$^{+}$ and
  Na$^{+}$ ions (see Supplementary Table \ref{SI-tab:diff}) }.  The
calculated and experimental $\xi_{\mathrm{max}}$ as a function of
$\lambda_{\mathrm{D}}$ are consequently plotted, for the seven salt
species considered here (\Fig{} \ref{fig:6}d).  \textcolor{blue}{
  Similar to the experimental observations, slightly higher $\xi$ for
  salts with same valence (e.g. LiCl, NaCl and KCl) is also observed
  in the simulations due to the variation in ionic mobilities as well
  as the possible specific surface adsorption of ions.} We notice that
without any fitting parameters, our theory can predict the
experimental values reasonably well.  In addition, the
$\xi_{\mathrm{max}}-\lambda_{\mathrm{D}}$ dependence is approximately
linear, suggesting that the effect of Debye length is dominant over
the salt diffusivity.
This
finding also explains the concentration dependence of
$\xi_{\mathrm{max}}$ observed in \Fig{} \ref{fig:2}d, as
$\lambda_{\mathrm{D}} \propto c_{0}^{-0.5}$.
\textcolor{blue}{It is worth noting that the
  observed linearity can only describe the $\lambda_{\mathrm{D}}$
  range considered here. When $\lambda_{\mathrm{D}}$ increases beyond
  30 nm, this linear relation would yield a $\xi_{\mathrm{max}}$
  value larger than 1, which appears to be unphysical.  More follow-up
  experimental and theoretical efforts are required to further
  investigate the small/large $\lambda_{\mathrm{D}}$ regimes. Also
  note that the salt diffusivity (see Eq. (4)) and the specific
  surface adsorption of ions may also influence the experimentally
  observed $\xi_{\mathrm{max}}$ beyond the simple
  $\lambda_{\mathrm{D}}$ effect.}
 Further including the
specific interaction of ionic species with the graphene membrane {-
  such as cation-$\pi$ interacation \cite{shi2013ion}} - may give even
better agreement with experimental results \cite{Ghosh_2018}. More
interestingly, based on our theoretical analysis, we predict that a
very high degree of salt rejection ($\xi>$ 0.99) can be already
achieved when $\lambda / r_{\mathrm{G}}>$3. In other words, for
example, when the Debye length is 20 nm, one can suppress 99 \% of the
diffusive flux by further reducing the pore radius down to $\sim$5-6
nm, which still remains larger than the hydrated radii. In combination
with the development of nanoscale etching techniques, we believe that
the physical picture presented here makes one step closer to
large-scale production of graphene membranes.

@@

\section*{Acknowledgment}
\label{sec:acknowledgment}

% \begin{acknowledgement}
  R.M.W. acknowledges Dr. Kyoungjun Choi from ETH Z\"{u}rich for the measurement of the electron mobility 
  of patterned graphene and is grateful for the support from the Binnig and Rohrer
  Nanotechnology Center of ETH Z\"{u}rich and IBM Z\"{u}rich. T.T. and
  C.J.S. are grateful for the financial support from the ETH startup
  funding. 
% \end{acknowledgement}

@@

\section{Supporting Information}
\label{sec:acknowledgment-1}
Detailed descriptions and discussions about graphene membrane
manufacturing, ion diffusion measurement, control transport
experiment with copper tape, circuit analog transport model, salt
Debye length and conductivity table, graphene charge carrier density
measurements, ionic pathway mechanism, effect of pore size
distribution, and implementation of numerical simulations

@@

\begin{figure}[H]
  \centering
  \includegraphics[width=0.95\linewidth]{img/fig2.pdf}
  \caption{ \textbf{The experimental evidence of salt rejection.}
    \textbf{a}. An example of measured conductivity in LCR as a
    function of time. The slope of conductivity-time curve when
    $V_{\mathrm{G}}>0$ is interpolated, is smaller than that without
    gating. \textbf{b}. Molar diffusive flux through bare PCTE (blue
    bars: experimental, $\boldsymbol{J}_{\mathrm{PCTE}}$ Exp.; red
    bars: model values, $\boldsymbol{J}_{\mathrm{PCTE}}$
    (Eq. \ref{eq:j-pcte})) and PG+PCTE
    ($\boldsymbol{J}_{\mathrm{PG0}}$ Exp.; green bars) membranes for
    different salts. \textcolor{blue}{The error bars for
      $\boldsymbol{J}_{\mathrm{PCTE}}$ Exp. were estimated by the
      standard deviation of three consecutive measurements. The error bars
      for $\boldsymbol{J}_{\mathrm{PCTE}}$ (Eq. \ref{eq:j-pcte})
      correspond to the calculated maximum and minimum ion flux values
      using the membrane parameters provided by the vendor. The error
      bars for $\boldsymbol{J}_{\mathrm{PG0}}$ Exp. were estimated by
      the standard deviation of the ion fluxes in the first interval
      during the gating experments.)}  \textbf{c}. Experimentally
    measured salt rejection factor $\xi$ as a function of
    $V_{\mathrm{G}}$ for 0.1 mM KCl solution in HCR, showing an
    asymmetric response, with a higher degree of salt rejection for
    $V_{\mathrm{G}}>0$. \textcolor{blue}{The error bars were estimated
      by the standard deviation for the ion fluxes in the second
      interval during the gating experiments.}
    \textbf{d}. $\xi_{\mathrm{max}}$ as a function of KCl
    concentration, fitted by a power law relation, suggesting the
    effect of Debye length on the observed salt rejection, since
    $\lambda_{\mathrm{D}} \propto c_{0}^{-\frac{1}{2}}$.}
  \label{fig:2} 
\end{figure}

@@

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End:
